EGGs are adhesive!

### 2025-07-03 jlamp/finalArxiv.tex
JLAMP

%\Copyright{Roberto Biondo and Davide Castelnovo and Fabio Gadducci} 
\iffalse
%TODO: questa roba negli argomenti
\ccsdesc[500]{Theory of Computation~Models of computation}
\ccsdesc[500]{Theory of Computation~Semantics and reasoning}

\keywords{Hypergraphs, terms graphs, e-graphs, adhesive categories.} %TODO mandatory; please add comma-separated list of keywords

\funding{The research has been partially supported by the Italian MUR - Call PRIN 2022 - Project 20228KXFN2 ``Spatio-Temporal Enhancement of Neural nets for Deeply
Hierarchical Automatised Logic'' (STENDHAL) 
and by the University of Pisa - Call PRA 2022 - Project 2022\_99 ``Formal Methods for the Healthcare Domain based on Spatial Information'' (FM4HD).}

%\nolinenumbers
\fi
\begin{document}

@@

\begin{abstract}
The use of rewriting-based visual formalisms is on the rise. 
%
In the formal methods community, this is due also to the introduction of adhesive
categories, where most properties of classical approaches to graph transformation, 
such as those on parallelism and confluence, can be rephrased and proved in a general and 
uniform way.
%
E-graphs (EGGs) are a formalism for program optimisation 
via an efficient implementation of equality saturation. 
In short, EGGs can be  defined as (acyclic) term graphs with an additional notion of 
equivalence on nodes that is closed under the operators of the signature.
Instead of replacing the components of a program, the optimisation step 
is performed by adding new components and linking them to 
the existing ones via an equivalence relation, until an optimal program is reached.
%
This work describes EGGs via adhesive categories. 
Besides the benefits in itself of a formal presentation, which renders the 
properties of the data structure precise, the description of the addition of equivalent 
program components using standard graph transformation tools offers the advantages 
of the adhesive framework in modelling, for example, concurrent updates.
%
%Besides 
%the benefits \emph{per se} of a formal presentation, making precise the properties of 
%the data structure, describing the addition of equivalent program components 
%via the standard tools of graph transformation offers the advantages given by the 
%adhesive framework in modelling e.g. concurrent updates.
\end{abstract}

>>>>

%\Copyright{Roberto Biondo and Davide Castelnovo and Fabio Gadducci}

\iffalse
%TODO: questa roba negli argomenti
\ccsdesc[500]{Theory of Computation~Models of computation}
\ccsdesc[500]{Theory of Computation~Semantics and reasoning}
\fi

\begin{document}

%\maketitle

\begin{abstract}[S U M M A R Y]
The use of rewriting-based visual formalisms is on the rise. 
%
In the formal methods community, this is due also to the introduction of adhesive
categories, where most properties of classical approaches to graph transformation, 
such as those on parallelism and confluence, can be rephrased and proved in a general and 
uniform way.
%
E-graphs (EGGs) are a formalism for program optimisation 
via an efficient implementation of equality saturation. 
In short, EGGs can be  defined as (acyclic) term graphs with an additional notion of 
equivalence on nodes that is closed under the operators of the signature.
Instead of replacing the components of a program, the optimisation step 
is performed by adding new components and linking them to 
the existing ones via an equivalence relation, until an optimal program is reached.
%
This work describes EGGs via adhesive categories. 
Besides the benefits in itself of a formal presentation, which renders the 
properties of the data structure precise, the description of the addition of equivalent 
program components using standard graph transformation tools offers the advantages 
of the adhesive framework in modelling, for example, concurrent updates.
%
%Besides 
%the benefits \emph{per se} of a formal presentation, making precise the properties of 
%the data structure, describing the addition of equivalent program components 
%via the standard tools of graph transformation offers the advantages given by the 
%adhesive framework in modelling e.g. concurrent updates.
\end{abstract}

@@
